comparing_scenario

### 2021-02-01 comparing_scenario.Rmd
Comparing original, weighted, and old data

```{r}
load("wages_hs2020.rda")
load("wages_hs2020_weighted.rda")
```

```{r}
bp_ori<- ggplot(wages_hs2020) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Original")

bp_wei <- ggplot(wages_hs2020_weighted) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Weighted")

bp_ori
bp_wei

@@

```{r}
wages_ori <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc1 <- as_tsibble(x = wages_hs2020_sc1,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc2 <- as_tsibble(x = wages_hs2020_sc2,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_weighted <- as_tsibble(x = wages_hs2020_weighted,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
lp_ori <- ggplot(wages_ori) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Original")

lp_sc1 <- ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 1")

lp_sc2 <- ggplot(wages_sc2) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 2")

lp_ori + lp_sc1 + lp_sc2

```
```{r}

get_obs <- function(data){
  data %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
  }

wages_obs_ori <- get_obs(wages_ori)
wages_obs_sc1 <- get_obs(wages_sc1)
wages_obs_sc2 <- get_obs(wages_sc2)
wages_obs_weig <- get_obs(wages_weighted)

nrow(wages_obs_ori)
nrow(wages_obs_sc1)
nrow(wages_obs_sc2)
nrow(wages_obs_weig)
```

```{r}
summary(wages_ori$mean_hourly_wage)
```

```{r}
summary(wages_weighted$mean_hourly_wage)
```

>>>>

```{r}
load("wages_hs2020.rda")
load("wages_hs2020_weighted.rda")
wages_old <- brolgar::wages %>%
  mutate(mean_hourly_wage = exp(ln_wages))
```

```{r}

# putting into tsibble
wages_ori <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_weighted <- as_tsibble(x = wages_hs2020_weighted,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_old <- as_tsibble(x = wages_old,
                    key = id,
                    index = xp,
                    regular = FALSE)
```

## The summary statistics

```{r}
get_summary <- function(data){
  summary(data$mean_hourly_wage)
}

print("Original data")
get_summary(wages_ori)

print("Weighted data")
get_summary(wages_weighted)

print("Old data")
get_summary(wages_old)

print("Weighted data, extreme value excluded")
get_summary(filter(wages_weighted, flag2 == "non_ext_val"))
```

## Number of ids in each data and number of observations per id

```{r}
get_obs <- function(data){
  data %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
}

wages_obs_ori <- get_obs(wages_ori)
wages_obs_weig <- get_obs(wages_weighted)
wages_obs_old <- get_obs(wages_old)

nrow(wages_obs_ori)
nrow(wages_obs_weig)
nrow(wages_obs_old)
```

```{r}

get_obs_plot <- function(df, plot_title){
  ggplot(df, aes(x = n_obs)) +
  geom_bar() +
  xlab("Number of observations") +
  ggtitle(plot_title)
}

get_obs_plot(wages_obs_ori, "Original data")
get_obs_plot(wages_obs_weig, "Weighted data")
get_obs_plot(wages_obs_old, "Old data")
```

## First time and last time participated in study (Original vs Weighted data)

```{r}
get_1st_time <- function(df, plot_title){
  df %>%
    features(year, list(min = min)) %>%
    ggplot(aes(x = min)) +
    geom_histogram(binwidth = 1) +
    xlab("First time in study") +
    ggtitle(plot_title)
}

get_1st_time(wages_ori, "Original data")
get_1st_time(wages_weighted, "Weighted data")
```

```{r}
get_last_time <- function(df, plot_title){
  df %>%
    features(year, list(max = max)) %>%
    ggplot(aes(x = max)) +
    geom_histogram(binwidth = 1) +
    xlab("First time in study") +
    ggtitle(plot_title)
}

get_last_time(wages_ori, "Original data")
get_last_time(wages_weighted, "Weighted data")
```

## The spaghetty plot

```{r}

get_sp <- function(df, plot_title, x){
  ggplot(df) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle(plot_title)
}

get_sp(wages_ori, "Original data", year)
get_sp(wages_weighted, "Weighted data", year)

ggplot(wages_old) +
  geom_line(aes(x = xp,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Old data")

@@

```{r}
get_sp(filter(wages_weighted, flag2 == "non_ext_val"), "Weighted data - extreme value excluded")
```

## Take a small spoon of those spaghetty

```{r}

set.seed(202102)

plot_sample <- function(df, plot_title){
  ggplot(df) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_sample() +
  ggtitle(plot_title)
}

plot_sample(wages_ori, "Original data")
plot_sample(wages_weighted, "Weighted data")

ggplot(wages_old) +
  geom_line(aes(x = xp,
                y = mean_hourly_wage,
                group = id)) +
  facet_sample() +
  ggtitle("Old data")
